---
title: "Expedition Explorer - Republic of the Marshall Islands 2023"
format: 
  dashboard:
    orientation: columns
    nav-buttons: [github]
    github: https://github.com/pristine-seas/exp-RMI-2023
logo: imgs/logo.png
theme: yeti
fig-width: 10
fig-asp: 0.3
self-contained: true
params:
  month: "September - October"
  year: "2023"
  # 2021 rates: https://www.cdc.gov/nchs/data/nvsr/nvsr72/nvsr72-01.pdf
  us_cesarean_rate: 0.321 
  us_preterm_rate:  0.1049
  threshold_diff: 0.02
editor_options: 
  chunk_output_type: console
---

```{r, message=F, warning = F, echo = F}
library(paletteer)
library(sf)
library(terra)
library(tidyverse)
library(plotly)

knitr::opts_chunk$set(eval = F, warning = F, message = F, include = F, echo = F)

PristineSeasR::set_ps_paths(email = "marine.data.science@ngs.org")

exp_path <- file.path(ps_science_path, "expeditions", "RMI-2023")

ps_data_path <- file.path(ps_science_path, "datasets/")

ggthemr::ggthemr('light')
```

```{r, eval = TRUE}
load(file.path(exp_path, "data/primary/output/expedition_metadata.Rdata"))

master_meta <- read_csv(file.path(exp_path, "data/primary/output/consolidated_expedition_metadata.csv"))

master_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"))

master_meta_shp <- sf::st_read(file.path(exp_path, "data/primary/processed/metadata/expedition_metadata.shp"))

argo_track <- sf::st_read(file.path(exp_path, "data/primary/raw/argo_route.geojson")) 

argo_nm <- round((st_length(argo_track)/10^3)*0.539957)

data_size <- "18 Tb"
```

#  {.sidebar}

This dashboard displays metadata for:

|             |                                 |
|-------------|---------------------------------|
| **Country** | Republic of the Marshall Island |
| **Year**    | September, 2023                 |
| **Vessel**  | R/V Argo                        |

::: {.callout-note collapse="true"}
## Disclaimer

This is a test dashboard.
:::

# All

## Column {width="65%"}

```{r eval = TRUE, include = T}
library(mapview)

master_meta_shp <- master_meta_shp %>% 
  distinct(method, ps_stt_, locatin, geometry)

mapview::mapview(master_meta_shp, 
                 zcol = c(method = "method"), 
                 legend = TRUE,
                 col.regions = paletteer_d("ggthemes::Tableau_10", 10),
                 map.types = c("Esri.WorldImagery"),
                 layer.name = "Methods",
                 popup = leafpop::popupTable(master_meta_shp,
                                      zcol = c("method",
                                               "ps_stt_",
                                               "locatin")))
```

## Column {width="35%"}

```{r, eval = TRUE, include = T}
master_meta |> 
  filter(location %in% c("Bikar", "Bokak", "Bikini", "Rongerik")) |> 
  group_by(location, method) |> 
  summarize(n_stations = n_distinct(ps_station_id)) |> 
  ungroup() |> 
  pivot_wider(names_from = "location", values_from = "n_stations", values_fill = 0) |> 
  mutate(Total = Bikar + Bikini + Bokak + Rongerik) |> 
  gt::gt(rowname_col = "method") |> 
  gt::tab_header(title = "Number of workstations in each island") |> 
  gt::grand_summary_rows(fns = "sum") |> 
  gt::tab_options(grand_summary_row.background.color = "lightblue")
```

### Row

#### Column

```{r eval = TRUE, include = T}
mapview::mapview(argo_track, 
                 map.types = c("Esri.WorldImagery"))
```

#### Column

```{r eval = T, include = T}
#| component: valuebox
#| title: Nautical miles traveled
list(
  icon = "rocket",
  color = "primary",
  value = as.numeric(argo_nm)
)
```

```{r eval = T, include = T}
#| component: valuebox
#| title: Hours underwater
list(
  icon = "water",
  color = "primary",
  value = 200
)
```

```{r eval = T, include = T}
#| component: valuebox
#| title: Data collected
list(
  icon = "floppy2",
  color = "primary",
  value = data_size
)
```

# Fish Surveys

```{r  eval = TRUE}
fish_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "fish")

fish_meta <- exp_meta$fish

load(file = file.path(exp_path, "data/primary/output/fish_surveys.Rdata"))

summary_by_strata <- biomass_by_strata_and_taxa |> 
  left_join(fish_meta |> distinct(ps_station_id, location, exposure, habitat)) |> 
  group_by(location, ps_station_id, depth_strata, depth_m, exposure, habitat) |> 
  summarise(n_taxa = n_distinct(taxon_valid_name[ind_m2 > 0]),
            across(c("ind_m2",  "gr_m2"), sum, na.rm = T)) |> 
  ungroup()

summary_by_taxa <- biomass_by_strata_and_taxa |> 
  left_join(fish_meta |> distinct(ps_station_id, location)) |> 
  group_by(location, taxon_valid_name) |>
  summarise(frq = 100*sum(gr_m2 > 0)/n(),
            across(c("ind_m2",  "gr_m2"), mean, na.rm = T)) |> 
  ungroup() |> 
  left_join(fish_taxa |> distinct(taxon_valid_name, common_family, trophic, family, class, order))
```

## Column {width="40%"}

### Column {.tabset}

```{r Bikar, eval = TRUE, include = T}
#| title: Bikar

fish_meta_sf |> 
  filter(location == "Bikar") |> 
  left_join(summary_by_strata |> select(ps_station_id, depth_strata, gr_m2)) |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("habitat"), 
                   legend = TRUE,
                   cex = c("gr_m2"),
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(fish_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))

summary_by_taxa |> 
  filter(location %in% c("Bikar")) |> 
  slice_max(gr_m2, n = 10) |> 
  mutate_if(is.numeric, round, 3) |> 
  select(-location) |> 
  left_join(fish_taxa |> distinct(taxon_valid_name, common_family)) |> 
  select(common_family, taxon_valid_name, frq, ind_m2, gr_m2) |> 
  gt::gt(rowname_col = "common_family") |> 
  gt::tab_header(title = "Reef fish with largest contribution to total biomass (g/m2)")
```

```{r Bokak, eval = TRUE, include = T}
#| title: Bokak
fish_meta_sf |> 
  filter(location == "Bokak") |> 
  left_join(summary_by_strata |> select(ps_station_id, depth_strata, gr_m2)) |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("habitat"), 
                   legend = TRUE,
                   cex = c("gr_m2"),
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(fish_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))


summary_by_taxa |> 
  filter(location %in% c("Bokak")) |> 
  slice_max(gr_m2, n = 10) |> 
  mutate_if(is.numeric, round, 3) |> 
  select(-location) |> 
  left_join(fish_taxa |> distinct(taxon_valid_name, common_family)) |> 
  select(common_family, taxon_valid_name, frq, ind_m2, gr_m2) |> 
  gt::gt(rowname_col = "common_family") |> 
  gt::tab_header(title = "Reef fish with largest contribution to total biomass (g/m2)")
```

```{r Bikini, eval = TRUE, include = T}
#| title: Bikini
fish_meta_sf |> 
  filter(location == "Bikini") |> 
  left_join(summary_by_strata |> select(ps_station_id, depth_strata, gr_m2)) |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("habitat"), 
                   legend = TRUE,
                   cex = c("gr_m2"),
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(fish_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))

summary_by_taxa |> 
  filter(location %in% c("Bikini")) |> 
  slice_max(gr_m2, n = 10) |> 
  mutate_if(is.numeric, round, 3) |> 
  select(-location) |> 
  left_join(fish_taxa |> distinct(taxon_valid_name, common_family)) |> 
  select(common_family, taxon_valid_name, frq, ind_m2, gr_m2) |> 
  gt::gt(rowname_col = "common_family") |> 
  gt::tab_header(title = "Reef fish with largest contribution to total biomass (g/m2)")
```

```{r Rongerik, eval = TRUE, include = T}
#| title: Rongerik
fish_meta_sf |> 
  filter(location == "Rongerik") |> 
  left_join(summary_by_strata |> select(ps_station_id, depth_strata, gr_m2)) |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("habitat"), 
                   legend = TRUE,
                   cex = c("gr_m2"),
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(fish_meta_sf |> 
                                                 filter(location == "Rongerik"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))

summary_by_taxa |> 
  filter(location %in% c("Rongerik")) |> 
  slice_max(gr_m2, n = 10) |> 
  mutate_if(is.numeric, round, 3) |> 
  select(-location) |> 
  left_join(fish_taxa |> distinct(taxon_valid_name, common_family)) |> 
  select(common_family, taxon_valid_name, frq, ind_m2, gr_m2) |> 
  gt::gt(rowname_col = "common_family") |> 
  gt::tab_header(title = "Reef fish with largest contribution to total biomass (g/m2)")
```

## Column {width="60%"}

```{r eval = TRUE, include = T}
summary_by_taxa |> 
  filter(location %in% c("Bikar", "Bokak")) |> 
  group_by(order, family, common_family) |> 
  summarize(n_taxa = n_distinct(taxon_valid_name[frq > 0]),
            across(c("ind_m2", "gr_m2"), 
                   sum, na.rm = T)) |> 
  filter(n_taxa > 0) |> 
  ungroup() |> 
  ggplot(aes(area = n_taxa, 
             label = paste(family, n_taxa, sep = "\n"),
             fill = order,
             subgroup = order)) +
  treemapify::geom_treemap(show.legend = F)+
  treemapify::geom_treemap_text(colour = "black", place = "middle", reflow = T, min.size = 3)+
  labs(fill = "",
       title = "Number of fish species by family. Color represents taxonomic order")+
  scale_fill_manual(values = paletteer_d("ggsci::default_igv"))
```

### Row {height="50%"}

```{r eval = TRUE, include = T}
refs <- tibble(location = c("Niue", "Flint", "Beveridge","Millenium", "Vostok"), 
               gr_m2 = c(80, 240, 270, 315, 412))

summary_by_trophic <- biomass_by_strata_and_taxa |> 
  left_join(fish_meta |> distinct(ps_station_id, location)) |> 
  left_join(fish_taxa |> distinct(taxon_valid_name, common_family, trophic)) |> 
  group_by(location, ps_station_id, depth_strata, trophic) |> 
  summarise(across(c("ind_m2",  "gr_m2"), sum, na.rm = T)) |> 
  ungroup() |> 
  group_by(location, depth_strata, trophic) |> 
  summarise(across(c("ind_m2",  "gr_m2"), mean, na.rm = T)) |> 
  ungroup() |> 
  #filter(location %in% c("Bikar", "Bokak")) |> 
  group_by(location) |> 
  mutate(pct_biomass = 100*gr_m2/sum(gr_m2)) |> 
  arrange(location, pct_biomass) |> 
  mutate(location = paste(location, depth_strata, sep = "\n"),
         trophic = factor(trophic)) 

tmp <- summary_by_trophic |> 
  bind_rows(refs) |> 
  ggplot()+
  geom_col(aes(fct_reorder(location, gr_m2, sum, na.rm = T), y = gr_m2), fill = "grey")+
  geom_col(data = summary_by_trophic,
           aes(location, y = gr_m2, fill = fct_reorder(trophic, gr_m2)))+
  labs(x = "", y = "Average fish biomass (g/m2)", fill ="") 

ggplotly(tmp)
```

# Benthic Surveys

```{r eval = TRUE}
benthic_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "benthos")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikar

benthic_meta_sf |> 
  filter(location == "Bikar") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "habitat"), 
                   legend = TRUE,
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(benthic_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bokak

benthic_meta_sf |> 
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "habitat"), 
                   legend = TRUE,
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(benthic_meta_sf |> 
                                                 filter(location == "Bokak"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bikini

benthic_meta_sf |> 
  filter(location == "Bikini") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "habitat"), 
                   legend = TRUE,
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(benthic_meta_sf |> 
                                                 filter(location == "Bikini"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Rongerik

benthic_meta_sf |> 
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "habitat"), 
                   legend = TRUE,
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(benthic_meta_sf |> 
                                                 filter(location == "Rongerik"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$benthos |> 
  group_by(location, exposure, habitat) |> 
  summarize(n_stations = n_distinct(ps_station_id)) |> 
  pivot_wider(names_from = location, values_from = n_stations) |> 
  gt::gt(rowname_col = "habitat") |> 
  gt::tab_header(title = "Number of benthos surveys by island, habitat, and exposure") 
```

### Row

```{r eval = TRUE, include = T}
exp_meta$benthos |> 
  distinct(location, ps_station_id, date, local_time, lat, lon, exposure, habitat, team_lead) |> 
  gt::gt(rowname_col = "location") |> 
  gt::tab_header(title = "Benthos survey locations") |> 
  gt::opt_interactive(use_compact_mode = T)
```

# eDNA

```{r eval = TRUE}
edna_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "eDNA")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikar

edna_meta_sf |> 
  filter(location == "Bikar") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "habitat"), 
                   legend = TRUE,
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(edna_meta_sf |> 
                                                 filter(location == "Bikar") ,
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bokak

edna_meta_sf |> 
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "habitat"), 
                   legend = TRUE,
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(edna_meta_sf |> 
                                                 filter(location == "Bokak") ,
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bikini

edna_meta_sf |> 
  filter(location == "Bikini") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "habitat"), 
                   legend = TRUE,
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(edna_meta_sf |> 
                                                 filter(location == "Bikini") ,
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Rongerik

edna_meta_sf |> 
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "habitat"), 
                   legend = TRUE,
                   col.regions = paletteer_d("ggthemes::Tableau_10", 3),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Habitat",
                   popup = leafpop::popupTable(edna_meta_sf |> 
                                                 filter(location == "Rongerik") ,
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$edna |> 
  mutate(strata = if_else(depth_m == 0, "Surface", "Bottom")) |> 
  group_by(location, strata) |> 
  summarize(n_liters = sum(n_replicates)*2) |> 
  pivot_wider(names_from = strata, values_from = n_liters) |> 
  ungroup() |> 
  gt::gt(rowname_col = "location") |> 
  gt::tab_header(title = "Water filtered (liters) for eDNA extraction in each island")
```

### Row

# Bruvs

```{r  eval = TRUE}
bruvs_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "bruvs")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikar
bruvs_meta_sf |> 
  filter(location == "Bikar") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("depth_m"), 
                   legend = TRUE,
                   col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(bruvs_meta_sf |> 
                                                 filter(location == "Bikar") ,
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bokak
bruvs_meta_sf |> 
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "depth_m"), 
                   legend = TRUE,
                   col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(bruvs_meta_sf |> 
                                                 filter(location == "Bokak") ,
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bikini
bruvs_meta_sf |> 
  filter(location == "Bikini") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "depth_m"), 
                   legend = TRUE,
                   col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(bruvs_meta_sf |> 
                                                 filter(location == "Bikini") ,
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Rongerik
bruvs_meta_sf |> 
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c(method = "depth_m"), 
                   legend = TRUE,
                   col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(bruvs_meta_sf |> 
                                                 filter(location == "Rongerik") ,
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$bruvs |> 
  group_by(location) |> 
  summarize(n_deployments = n_distinct(ps_station_id),
            min_depth = round(min(depth_m)),
            max_depth = round(max(depth_m)),
            depth_range = paste(min_depth, max_depth, sep = "-")) |> 
  ungroup() |> 
  select(location, depth_range, n_deployments) |> 
  gt::gt(rowname_col = "location") |> 
  gt::tab_header(title = "Number of benthic BRUVS deployed in each island")
```

### Row

# Pelagics

```{r  eval = TRUE}
pelagics_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "pelagics")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikar
pelagics_meta_sf |> 
  filter(location == "Bikar") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   popup = leafpop::popupTable(pelagics_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bokak
pelagics_meta_sf |> 
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   popup = leafpop::popupTable(pelagics_meta_sf |> 
                                                 filter(location == "Bokak"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bikini
pelagics_meta_sf |> 
  filter(location == "Bikini") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   popup = leafpop::popupTable(pelagics_meta_sf |> 
                                                 filter(location == "Bikini"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Rongerik
pelagics_meta_sf |> 
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   popup = leafpop::popupTable(pelagics_meta_sf |> 
                                                 filter(location == "Rongerik"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "depth_m")))
```

## Column {width="30%"}

### Row

# Dropcams

```{r  eval = TRUE}
dscm_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "dscm")


dscm_meta_sf |> 
  filter(location == "Bikar") |> 
  ggplot()+
  geom_sf(aes(col = depth_m))
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikar
dscm_meta_sf |> 
  filter(location == "Bikar") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("depth_m"), 
                   legend = TRUE,
                   #col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(dscm_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date_in", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bokak
dscm_meta_sf |> 
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("depth_m"), 
                   legend = TRUE,
                   #col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(dscm_meta_sf |> 
                                                 filter(location == "Bokak"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date_in", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Bikini
dscm_meta_sf |> 
  filter(location == "Bikini") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("depth_m"), 
                   legend = TRUE,
                   #col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(dscm_meta_sf |> 
                                                 filter(location == "Bikini"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date_in", 
                                                        "depth_m")))
```

```{r eval = TRUE, include = T}
#| title: Rongerik
dscm_meta_sf |> 
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("depth_m"), 
                   legend = TRUE,
                   #col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(dscm_meta_sf |> 
                                                 filter(location == "Rongerik"),
                                               zcol = c("method",
                                                        "habitat",
                                                        "ps_station_id",
                                                        "location",
                                                        "date_in", 
                                                        "depth_m")))
```

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$dscm |> 
  mutate(depth_cut = cut(depth_m, 
                         breaks = c(0,200,500,1000,2500),
                         labels = c("0 - 200 m", "200 - 500 m", "500 - 1,000 m", "1,000 - 2,500 m"))) |> 
  group_by(location, depth_cut) |> 
  summarize(n_deployments = n_distinct(ps_station_id)) |> 
  ungroup() |> 
  pivot_wider(names_from = location, values_from = n_deployments) |> 
  gt::gt(rowname_col = "depth_cut") |> 
  gt::tab_header(title = "Number of deep sea cameras deployed in each island by depth strata")
```

### Row

# Sub

```{r  eval = TRUE}
sub_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "sub")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikar
sub_meta_sf |> 
  filter(location == "Bikar") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("max_depth_m"), 
                   legend = TRUE,
                   col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(sub_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "max_depth_m",
                                                        "max_depth_temp", 
                                                        "observer_1",
                                                        "observer_2"
                                                        )))
```

```{r eval = TRUE, include = T}
#| title: Bokak
sub_meta_sf |> 
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("max_depth_m"), 
                   legend = TRUE,
                   col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(sub_meta_sf |> 
                                                 filter(location == "Bokak"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "max_depth_m",
                                                        "max_depth_temp", 
                                                        "observer_1",
                                                        "observer_2"
                                                        )))
```

```{r eval = TRUE, include = T}
#| title: Bikini
sub_meta_sf |> 
  filter(location == "Bikini") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("max_depth_m"), 
                   legend = TRUE,
                   col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(sub_meta_sf |> 
                                                 filter(location == "Bikini"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "max_depth_m",
                                                        "max_depth_temp", 
                                                        "observer_1",
                                                        "observer_2"
                                                        )))
```

```{r eval = TRUE, include = T}
#| title: Rongerik
sub_meta_sf |> 
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(zcol = c("max_depth_m"), 
                   legend = TRUE,
                   col.regions = paletteer_c("grDevices::Viridis", 5),
                   map.types = c("Esri.WorldImagery"),
                   layer.name = "Depth (m)",
                   popup = leafpop::popupTable(sub_meta_sf |> 
                                                 filter(location == "Rongerik"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "max_depth_m",
                                                        "max_depth_temp", 
                                                        "observer_1",
                                                        "observer_2"
                                                        )))
```

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$sub |> 
  group_by(location) |> 
  summarize(n_deployments = n_distinct(ps_station_id),
            min_depth = min(max_depth_m),
            max_depth = max(max_depth_m)) |> 
  ungroup() |> 
  gt::gt(rowname_col = "location") |> 
  gt::tab_header(title = "Number of submersible surveys in each island")
```

### Row

# Birds

```{r  eval = TRUE}
birds_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "birds")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikar
birds_meta_sf |>  
  filter(location == "Bikar") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   col.regions = "yellow",
                   col = "yellow",
                   popup = leafpop::popupTable(birds_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "vessel")))
```

```{r eval = TRUE, include = T}
#| title: Bokak
birds_meta_sf |>  
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   col.regions = "yellow",
                   col = "yellow",
                   popup = leafpop::popupTable(birds_meta_sf |> 
                                                 filter(location == "Bokak"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "vessel")))
```

```{r eval = TRUE, include = T}
#| title: Bikini
birds_meta_sf |>  
  filter(location == "Bikini") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   col.regions = "yellow",
                   col = "yellow",
                   popup = leafpop::popupTable(birds_meta_sf |> 
                                                 filter(location == "Bikini"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "vessel")))
```

```{r eval = TRUE, include = T}
#| title: Rongerik
birds_meta_sf |>  
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   col.regions = "yellow",
                   col = "yellow",
                   popup = leafpop::popupTable(birds_meta_sf |> 
                                                 filter(location == "Rongerik"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date", 
                                                        "vessel")))
```

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$birds |>
  mutate(type = if_else(vessel == "Land", "Land", "Sea")) |> 
  group_by(location, type) |> 
  summarize(n_surveys = n_distinct(ps_station_id)) |> 
  pivot_wider(names_from = type, values_from = n_surveys) |> 
  ungroup() |> 
  gt::gt(rowname_col = "location") |> 
  gt::tab_header(title = "Number of bird surveys in each island")
```

### Row

# Vegetation

```{r  eval = TRUE}
veg_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "vegetation")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikar
veg_meta_sf |>  
  filter(location == "Bikar") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   popup = leafpop::popupTable(veg_meta_sf |> 
                                                 filter(location == "Bikar"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "survey_type",
                                                        "size_m",
                                                        "team_lead")))
```

```{r eval = TRUE, include = T}
#| title: Bokak
veg_meta_sf |>  
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   popup = leafpop::popupTable(veg_meta_sf |> 
                                                 filter(location == "Bokak"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "survey_type",
                                                        "size_m",
                                                        "team_lead")))
```

```{r eval = TRUE, include = T}
#| title: Rongerik
veg_meta_sf |>  
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   legend = FALSE,
                   popup = leafpop::popupTable(veg_meta_sf |> 
                                                 filter(location == "Rongerik"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "survey_type",
                                                        "size_m",
                                                        "team_lead")))
```

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$vegetation |>
  group_by(location, survey_type) |> 
  summarize(n_surveys = n_distinct(ps_station_id)) |> 
  pivot_wider(names_from = survey_type, values_from = n_surveys) |> 
  ungroup() |> 
  gt::gt(rowname_col = "location") |> 
  gt::tab_header(title = "Number of vegetation surveys per island")
```

### Row

# Traverses

```{r  eval = TRUE}
traverse_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "traverses")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bikini
traverse_meta_sf |>  
  filter(location == "Bikini") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   col.regions = "pink",
                   col = "pink",
                   legend = FALSE,
                   popup = leafpop::popupTable(traverse_meta_sf |> 
                                                 filter(location == "Bikini") ,
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "USGS_traverse_id",
                                                        "distance_m",
                                                        "team_lead")))
```

```{r eval = TRUE, include = T}
#| title: Rongerik
traverse_meta_sf |>  
  filter(location == "Rongerik") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(map.types = c("Esri.WorldImagery"),
                   col.regions = "pink",
                   col = "pink",
                   legend = FALSE,
                   popup = leafpop::popupTable(traverse_meta_sf |> 
                                                 filter(location == "Rongerik") ,
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "USGS_traverse_id",
                                                        "distance_m",
                                                        "team_lead")))
```

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$traverses |>
  group_by(location) |> 
  summarize(n_surveys = n_distinct(ps_station_id),
            cum_dist_m = sum(distance_m)) |> 
  ungroup() |> 
  gt::gt(rowname_col = "location") |> 
  gt::tab_header(title = "Number of traverses and distance covered (m2) per island")
```

### Row

# Plastics

```{r  eval = TRUE}
plastics_meta_sf <- sf::st_read(file.path(exp_path, "data/primary/output/expedition_metadata.gpkg"), 
                            layer = "plastics")
```

## Column {width="70%"}

### Column {.tabset}

```{r eval = TRUE, include = T}
#| title: Bokak
plastics_meta_sf |>  
  filter(location == "Bokak") |> 
  st_jitter(factor = 0.001) |> 
  mapview::mapview(legend = FALSE,
                   map.types = c("Esri.WorldImagery"),
                   popup = leafpop::popupTable(plastics_meta_sf |> 
                                                 filter(location == "Bokak"),
                                               zcol = c("method",
                                                        "ps_station_id",
                                                        "location",
                                                        "date",
                                                        "team_lead")))
```

## Column {width="30%"}

### Row

# Turtles

## Column {width="70%"}

### Column {.tabset}

## Column {width="30%"}

```{r eval = TRUE, include = T}
exp_meta$turtles |> 
  group_by(location, sublocation) |> 
  summarize(n_surveys = n()) |> 
    ungroup() |> 
  gt::gt(rowname_col = "location") |> 
  gt::tab_header(title = "Number of turle surveys per island")
```

### Row
